---
title: "TidyTuesday: California Forest Fires"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Analyzing data for #tidytuesday week of 8/21/2018

source: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-08-21

```{r}
# load packages and parse data
library(tidyverse)
library(scales)
library(RColorBrewer)
library(forcats)
library(ggmap)

theme_set(theme_bw())

calfire_incidents_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-08-21/cal-fire-incidents.csv")
calfire_incidents <- calfire_incidents_raw %>%
  select(year = YEAR,
         fires = `NUMBER OF FIRES`,
         acres_burned = `ACRES BURNED`,
         dollar_damage = `DOLLAR DAMAGE`)
inflation <- read_csv("adjust_inflation.csv")

calfire_frap_raw <- read_csv("https://github.com/rfordatascience/tidytuesday/raw/master/data/2018-08-21/week21_calfire_frap.csv")
calfire_frap <- calfire_frap_raw
```

### Total number of fires over time
Recent trends indicate this has steadily been on the decline over the years

```{r}
calfire_incidents %>% 
  ggplot(aes(year, fires)) + 
  geom_bar(stat = 'identity',
           fill = "#de2d26") +
  scale_y_continuous(labels = comma_format()) +
  labs(x = "",
       y = "")
```
Note: missing data for year of 1976

### Damage over time
Increase in dollar damage over time may suggest 1) where the fires are actually happening and 2) if more people are living in more fire-prone areas of California
```{r}
calfire_incidents %>%
  inner_join(inflation) %>%
  mutate(inflation_dmg = `2016_damage` / acres_burned) %>% 
  ggplot() +
  geom_bar(aes(year, inflation_dmg), 
           stat = 'identity',
           fill = '#de2d26') +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "",
       y = "",
       title = "$ Damage by Acres per year",
       subtitle = "Adjusted for inflation in 2016 $")
```

### What are the primary causes of forest fires?
```{r}
# make "cause coding" table
cause <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19)
description <- c('Lightning','Equipment Use','Smoking','Campfire','Debris','Railroad','Arson','Playing with Fire','Miscellaneous','Vehicle','Power Line','Firefighter Training','Non-Firefighter Training','Unknown/Unidentified','Structure','Aircraft','Volcanic','Escaped Prescribed Burn','Illegal Alien Campfire')
cause_coding <- as_tibble(cbind(cause, description))

# convert original variable from int to character to allow for joining of data
calfire_frap$cause <-as.character(calfire_frap$cause)

calfire_frap %>%
  inner_join(cause_coding) %>%
  count(description, sort = TRUE) %>%
  mutate(description = fct_reorder(description, n)) %>%
  ggplot(aes(description, n/sum(n))) + 
  geom_col(aes(fill = description)) + 
  coord_flip() +
  theme(legend.position = 'none') 



calfire_frap %>%
  inner_join(cause_coding) %>%
  count(year_, description, sort = TRUE) %>%
  ggplot(aes(year_, n/sum(n))) + 
  geom_line(aes(color = description)) + 
  theme(legend.position = 'none') 
```

### Which agency handles the most fires over the years?
```{r}
calfire_frap %>%
  select(year_, agency) %>%
  group_by(year_, agency) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(year_, n)) +
  geom_line(aes(color = agency))

```

```{r}

```

