---
title: "TidyTuesday: Thanksgiving Dinner"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Analyzing data for #tidytuesday week of 11/20/2018

source: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-11-20

```{r}
# load packages and parse data
library(tidyverse)
library(scales)
library(RColorBrewer)
library(forcats)

theme_set(theme_bw())

thanksgiving_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-11-20/thanksgiving_meals.csv")

thanksgiving <- 
  thanksgiving_raw %>%
  filter(celebrate != 'No')
```


### What are the most popular pies for Thanksgiving ?
```{r}
thanksgiving %>%
  select(pie1:pie13, age, gender) %>%
  gather(pie_type = pie1:pie13, 
         value = count) %>%
  select(pie_type = key, 
         pie = count) %>%
  filter(pie != 'NA', !grepl('Other', pie), pie != 'None') %>% 
  group_by(pie) %>%
  count() %>% 
  filter(n > 10) %>%
  ungroup() %>%
  ggplot(aes(reorder(pie, n), n, label = n)) +
  geom_bar(aes(fill = pie),
           alpha = 0.9,
           stat='identity') +
  geom_text(aes(label = n, color = pie), hjust = -0.5) +
  coord_flip() +
  theme_classic() + 
  theme(legend.position = 'none') +
  xlab("") +
  ylab("") +
  labs(title = "Most Popular Pies for Thanksgiving (n=980)",
        subtitle = "Question: Which type of pie is typically served at your Thanksgiving dinner? Please select all that apply")
```

### Most popular pies distributed by gender ?
```{r}
thanksgiving %>%
  select(pie1:pie13, age, gender) %>%
  gather(pie_type = pie1:pie13, 
         value = count) %>%
  select(gender, 
         pie_type = key,
         pie = count) %>%
  na.omit() %>%
  filter(!grepl('Other', pie), pie != 'None') %>% 
  group_by(gender, pie) %>%
  count() %>%
  filter(n > 10) %>%
  ungroup() %>%
  ggplot(aes(reorder(pie, n), n)) +
  geom_bar(stat = 'identity',
           position = 'fill',
           alpha = 0.9,
           aes(fill = gender)) + 
  scale_fill_brewer(palette = 'Set2') +
  coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "Most Popular Pies for Thanksgiving by Gender (n=980)") +
  scale_y_continuous(labels = percent_format())
```

### Most popular pies distributed by age group?
```{r}
thanksgiving %>%
  select(pie1:pie13, age, gender) %>%
  gather(pie_type = pie1:pie13, 
         value = count) %>%
  select(age, 
         pie_type = key,
         pie = count) %>%
  na.omit() %>%
  filter(!grepl('Other', pie), pie != 'None') %>% 
  group_by(age, pie) %>%
  count() %>%
  filter(n > 10) %>%
  ungroup() %>%
  ggplot(aes(reorder(pie, n), n)) +
  geom_bar(stat = 'identity',
           position = 'fill',
           alpha = 0.8,
           aes(fill = age)) + 
  scale_fill_brewer(palette = 'Set1') +
  coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "Most Popular Pies for Thanksgiving by Age Group (n=980)") +
  scale_y_continuous(labels = percent_format())
```




### For a given age group, what do they think is the cutoff age for the kids table?
```{r}
thanksgiving %>%
  select(age, kids_table_age) %>%
  na.omit() %>%
  ggplot() +
  geom_bar(aes(fct_rev(age), fill = kids_table_age), 
           position = 'fill',
           alpha = 0.9) +
  coord_flip() +
  scale_fill_brewer(palette = 'Paired') +
  xlab("") +
  ylab("") + 
  labs(title = "Graph title",
       subtitle = "Question: What is the age cutoff at your 'kids table' at Thanksgiving?") +
  scale_y_continuous(labels = percent_format())
```

### How do you cook your main dish?
```{r}
thanksgiving %>%
  select(main_dish, main_prep, stuffing, cranberry) %>%
  na.omit() %>%
  filter(!grepl('know|Other', main_dish)) %>%
  ggplot() + 
  geom_bar(aes(fct_rev(main_dish), fill = main_prep), 
           position = 'fill') + 
  coord_flip() + 
  scale_fill_brewer(palette = 'Paired') +
  scale_y_continuous(labels = percent_format()) +
  xlab("") +
  ylab("") +
  labs(title = "Most Popular Methods for Preparing Thanksgiving Dish",
       subtitle = "Question: How is the main dish typically cooked?")
```

### Further exploration available to check out the top sides for Thanksgiving dinner!
```{r}

```

